toyData

### 2017-11-09 latex/article_vol3.tex
Updated paper

A number of \proglang{R} packages made for other pre-analysis steps
are already available, including \pkg{janitor} \citep{janitor},
\pkg{assertive} \citep{assertive}, \pkg{dplyr} \citep{dplyr},
\pkg{tidyr}\citep{tidyr}, \pkg{data.table} \citep{data.table},
\pkg{DataCombine} \citep{DataCombine}, \pkg{validate}
\citep{validate}, \pkg{assertr}, and \citep{assertr}. These packages
focus on different stages of the pre-analysis work.  \pkg{janitor}
provides tools for data import with a particular emphasis on the
challenges of getting neat data frames from Microsoft Excel data
files. \pkg{dplyr}, \pkg{tidyr}, \pkg{data.table} and
\pkg{DataCombine} go a few steps further by providing a wide array of
extremely powerful tools for data wrangling, including a number of
particularly useful functions for merging and working with very large
datasets. When it comes to actual data cleaning, however, the options
are fewer.  \pkg{validate} (and the similar packages \pkg{editrules}
\citep{editrules} and \pkg{deducorrect} \citep{deducorrect} from the
same authors) and \pkg{assertive} offers tools for identifying errors
in a dataset by checking the state of the variable given a set of
pre-specified rules, and their focus is on internal validity rather
than general data screening. In practice, this means that quite elegant
tools for, e.g., linear restraints among the variables in a dataset
can be applied, but looking for potentially miscoded missing values is
not really feasible. The main difference between these two challenges
is the direction in which the data is inspected: While linear
constraints work observation-wise with no ambiguity, determining
whether or not something is a miscoded missing value often requires
knowledge about the full variable (e.g. range or data type), and thus
it should be performed variable-wise. \pkg{validate} does not
currently allow for user-defined extensions of the latter type,
thereby limiting its data cleaning potential.  Automatic data
correction functions are also provided by \pkg{validate} which we
consider to be quite a dangerous cocktail: all power is given to the
the computer with no human supervision, and investigators are less
likely to make an active, case-specific choice regarding the handling
of the potential errors.  Finally, no tools have been made to easily
document exactly which checks and preliminary results were used in the
data cleaning process. The \pkg{assertr} package provide very similar
--- and very nice --- tools to those of \pkg{validate}, but without
any amibitions of conducting auto-cleaning.

@@

By default, an \proglang{R} markdown file and a rendered pdf \hl{OBS: måske html/doc alt efter hvem der kalder funktionen?} overview
document is produced, saved to the working directory and
opened for immediate inspection. Such a data report consists of three parts, two of which 
are presented in Figure~\ref{fig:example1}. First, an overview of what was done is presented under the
title \textit{Data report overview}. Secondly, an index listing each variable along with an indication of whether it was found to be problematic or not is provided. Thirdly, as seen in Figure~\ref{fig:example2}, each variable in the
dataset is presented in turn using (up to) three tools in the
\textit{Variable list}: A table summarizing key features of the
variable, a figure visualizing its distribution and a
list of flagged issues, if any. For instance, as shown in Figure~\ref{fig:example2}, for the \code{numeric}-type variable
\code{events} from \code{toyData}, \code{makeDataReport()} has identified two values that
are suspected to be miscoded missing values (\code{999} and \code{NaN}),
while two values were also flagged as potential outliers that should
be investigated more carefully.

Though the \code{makeDataReport()} function is very easy to use, it should not be
mistaken to be inflexible: By using the optional arguments of the function, 
both the contents and the look of the data report can be molded 
according to the user's needs. The most commonly used arguments are summarized in
Table~\ref{table.cleanFormals} and they are grouped according to the
part of the data assesment and report generation they influence. In order to understand
this distinction, a glimpse of the inner structure of \code{makeDataReport()} is
shown in Figure~\ref{figure:cleanStructure}. Below, we present a few
examples on how to use the arguments from Table \ref{table.cleanFormals}
 to influence the output of a \code{makeDataReport()} call.

@@

\begin{Schunk}
\begin{Sinput}
R> makeDataReport(toyData, output="html", render=FALSE, openResult=FALSE, replace=TRUE)
\end{Sinput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
R> makeDataReport(toyData, preChecks = c("isKey", "isSupported"), replace=TRUE)
\end{Sinput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
R> makeDataReport(toyData, visuals = setVisuals(all = "basicVisual"), replace=TRUE)
\end{Sinput}
\end{Schunk}

In summary, and as indicated in Figure~\ref{figure:cleanStructure}, there are two stages
where \code{makeDataReport()} applies functions to each of the variables:
\begin{enumerate}
\item In the pre-check stage.
\item As part of the summarize/visualize/check (SVC) steps.
\end{enumerate}
Each of these stages are controllable using appropriate function
arguments in \code{makeDataReport()}, and above we have shown examples of how
to tweak them to modify the data cleaning outputs. However, if the
dataset at hand requires new, additional checks, then more control is
needed \hl{ and we have provided a detailed vignette that explains how to modify and
expand the possibilities by producing new summary, visual, and check
functions in [REFERENCE]}.

@@

\begin{Schunk}
\begin{Sinput}
R> complexData <- data.frame(complexVar = complex(100, real = 1:100, imaginary = 3), numericVar = 1:100)
R> makeDataReport(complexData, treatXasY=list(complex = "numeric"), replace=TRUE)
\end{Sinput}
\end{Schunk}

@@

\begin{figure}[tb]
\begin{center}
\includegraphics[width=7.5cm]{toyData-events.pdf}
\end{center}
\caption{Output from running \code{visualize()} on the variable \code{events} from the
\code{toyData} dataset.}
\label{fig:example3}
\end{figure}

@@

\begin{figure}[tb]
\begin{center}
\frame{\includegraphics[width=7.5cm,page=3]{dataMaid_presidentData.pdf}}
\frame{\includegraphics[width=7.5cm,page=4]{dataMaid_presidentData.pdf}}
\end{center}
\caption{The second and third pages of the \code{presidentData} data report. We see that there are two \code{Name} variables in the overview on page 2 and see that these variables are indeed treated as \code{character} variables on page 3, as specified in the \code{makeDataReport} call by use of the \code{treatXasY} argument. Larger versions of the pages can be seen in
  Appendix~\ref{sec:appendix2}}
\label{fig:bigExampleP23}
\end{figure}

\begin{figure}[tb]
\begin{center}
\frame{\includegraphics[width=7.5cm,page=5]{dataMaid_presidentData.pdf}}
\frame{\includegraphics[width=7.5cm,page=6]{dataMaid_presidentData.pdf}}
\end{center}
\caption{Part of the data overview report generated for the \code{presidentData} dataset. Here, we see pages 4 and 5 from the \textit{Variable list}. Larger versions of the pages can be seen in
  Appendix~\ref{sec:appendix2}}
\label{fig:bigExampleP45}
\end{figure}

We will now put the bits and pieces from above together and show how \code{makeDataReport()} can be used on a less artificial dataset to create a useful overview report and how the interactive tools can subsequently be used to assist the actual data cleaning process. More specifically, we will create a report describing the \code{presidentData} dataset, which is available in \pkg{dataMaid} and use the information from this report to clean up the data. \code{presidentData} is a sligtly mutilated dataset with information about the 45 first US presidents, but with a few common data issues and a blind passenger. The dataset contains one observation per president and has the following variables:
\begin{description}
\item[\code{lastName}] The last name of the president.
\item[\code{firstName}] The first name of the president.
\item[\code{orderOfPresidency}] The number in the order of presidents.
\item[\code{birthday}] The birthday of the president.
\item[\code{stateOfBirth}] The state in which the president was born.
\item[\code{assassinationAttempt}] Was there an assassination attempt on the president?
\item[\code{sex}] The sex of the president.
\item[\code{ethnicity}] The ethnicity of the president.
\item[\code{presidencyYears}] The duration of the presidency.
\item[\code{ageAtInauguration}] The age of the president at inauguration.
\item[\code{favoriteNumber}] The favourite number of the president (fictional).
\end{description}
We discuss the results of a data overview report generated for this dataset below, but first there are a few special features of the dataset and wishes for the data report that require us to customize it using some of the arguments for \code{makeDataReport}. We have the following points of interest:
\begin{enumerate}
\item A couple of variables are used to store names, namely \code{lastName} and \code{firstName}. In order to use special bibliographical analysis tools on these, and only these, variables, it might be convenient to assign them a special class. Therefore, these variables have been set to have class \code{Name} by use of the base \proglang{R} function \code{class()}. When we wish to make a data report for the dataset, we have to tell \code{makeDataReport()} how to handle such \code{Name}-type variables, as they are not among the supported variable types mentioned in the above. This can be done using the \code{treatXasY} argument.
\item We use the \code{character} class for a few categorical variables that have a lot of different levels, but where it is not a data mistake that these levels each generally have very few observations, e.g. the variable \code{stateOfBirth}. Therefore, we would like to disable the \code{identifyLoners} check (which flags variables with <6 observations in any of the levels) for this variable type. This can be done using the \code{checks} argument.
\item We would like the report to be called "Dirty president data" in order to reflect that it is, indeed, a report concerning dirty data about presidents.
\end{enumerate}
Incorporating these three customizations, we can load the data and generate a report by calling:

\begin{Schunk}
\begin{Sinput}
R> data(presidentData)
R> makeDataReport(presidentData, replace = TRUE, 
+    treatXasY = list("Name" = "character"),
+    checks = setChecks(character = 
+    defaultCharacterChecks(remove = "identifyLoners")),
+    reportTitle = "Dirty president data")
\end{Sinput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
R> presidentData$lastName[presidentData$lastName == " Truman"] <- "Truman"
R> presidentData$stateOfBirth[presidentData$stateOfBirth == "New york"] <- "New York"
R> presidentData$assassinationAttempt <- factor(presidentData$assassinationAttempt)
R> presidentData$ageAtInauguration <- as.numeric(presidentData$ageAtInauguration)
\end{Sinput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
R> presidentData$ageAtInauguration <- as.numeric(as.character(presidentData$ageAtInauguration))
\end{Sinput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
R> birthdayOutlierVal <- 
+  +    identifyOutliers(presidentData$birthday)$problemValues
\end{Sinput}
\end{Schunk}

@@

\subsection[Incorporating dataMaid in a programmer's workflow]{Incorporating \pkg{dataMaid} in a programmer's workflow}
The default settings of \code{makeDataReport()} have been set to facilitate easy and quick data report generation, but unfortunately, this also means that it is not ideal for a more programming-oriented workflow, where the function might not be called by a human. For instance, one might be interested in automatically running \code{makeDataReport()} on all datasets received from a certain client, perhaps via email or through a web upload, and returning a data report for them to inspect and comment before ever looking at the data. In this scenario, there are a few issues with the standard data report:
\begin{enumerate}
\item After rendering, the report is automatically opened. This is not very useful, if the processes are supposed to run in the background.
\item The report generation writes messages to the console while producing and rendering the report.
\item Unless specifically told otherwise, every report created for the same dataset (or different datasets with the same storage name in \proglang{R}) will have the same file name.
\end{enumerate}
Please note that the data report does contain information about who, when and how concerning its generation, so even though the default choices for file names do not make it easy to tell different reports for the same dataset apart, it should be rather easy when inspecting the report manually. These three problems can easily be solved by use of the arguments of \code{makeDataReport()}. Whether or not the outputted file is opened can be controlled through the argument \code{open}. How much information is printed in the console can be adjusted by using the argument \code{quiet}. And convenintely introducing small alterations of the file names can be obtained by use of the \code{vol} argument. For instance, we can make a data report for \code{toyData} that is not opened automatically, produces no output to the console and includes the date and time of its creation in the file name:

\begin{Schunk}
\begin{Sinput}
>R makeDataReport(toyData, open = FALSE, quiet = "silent",
+   vol = paste("_", format(Sys.time(), "%m-%d%-%y_%H.%M"), sep = ""))
\end{Sinput}
\end{Schunk}

>>>>

A number of \proglang{R} packages made for other pre-analysis steps
are already available, including \pkg{janitor} \citep{janitor},
\pkg{assertive} \citep{assertive}, \pkg{dplyr} \citep{dplyr},
\pkg{tidyr}\citep{tidyr}, \pkg{data.table} \citep{data.table},
\pkg{DataCombine} \citep{DataCombine}, \pkg{validate}
\citep{validate}, and \pkg{assertr} \citep{assertr}. These packages
focus on different stages of the pre-analysis work.  \pkg{janitor}
provides tools for data import with a particular emphasis on the
challenges of getting neat data frames from Microsoft Excel data
files. \pkg{dplyr}, \pkg{tidyr}, \pkg{data.table} and
\pkg{DataCombine} go a few steps further by providing a wide array of
extremely powerful tools for data wrangling, including a number of
particularly useful functions for merging and working with very large
datasets. When it comes to actual data cleaning, however, the options
are fewer.  \pkg{validate} (and the similar packages \pkg{editrules}
\citep{editrules} and \pkg{deducorrect} \citep{deducorrect} from the
same authors) and \pkg{assertive} offers tools for identifying errors
in a dataset by checking the state of the variable given a set of
pre-specified rules, and their focus is on internal validity rather
than general data screening. In practice, this means that quite elegant
tools for, e.g., linear restraints among the variables in a dataset
can be applied, but looking for potentially miscoded missing values is
not really feasible. The main difference between these two challenges
is the direction in which the data is inspected: While linear
constraints work observation-wise with no ambiguity, determining
whether or not something is a miscoded missing value often requires
knowledge about the full variable (e.g. range or data type), and thus
it should be performed variable-wise. \pkg{validate} does not
currently allow for user-defined extensions of the latter type,
thereby limiting its data cleaning potential.  Automatic data
correction functions are also provided by \pkg{validate} which we
consider to be quite a dangerous cocktail: all power is given to the
the computer with no human supervision, and investigators are less
likely to make an active, case-specific choice regarding the handling
of the potential errors.  Finally, no tools have been made to easily
document exactly which checks and preliminary results were used in the
data cleaning process. The \pkg{assertr} package provide very similar
--- and very nice --- tools to those of \pkg{validate}, but without
any amibitions of conducting auto-cleaning.

@@

By default, an \proglang{R} markdown file and a rendered pdf or word
overview document is produced, saved to the working directory and
opened for immediate inspection. Such a data report consists of three
parts, two of which are presented in Figure~\ref{fig:example1}. First,
an overview of what was done is presented under the title \textit{Data
  report overview}. Secondly, an index listing each variable along
with an indication of whether it was found to be problematic or not is
provided. Thirdly, as seen in Figure~\ref{fig:example2}, each variable
in the dataset is presented in turn using (up to) three tools in the
\textit{Variable list}: A table summarizing key features of the
variable, a figure visualizing its distribution when relevant and a list of flagged
issues, if any. For instance, as shown in Figure~\ref{fig:example2},
for the \code{numeric}-type variable \code{events} from
\code{toyData}, \code{makeDataReport()} has identified two values that
are suspected to be miscoded missing values (\code{999} and
\code{NaN}), while two values were also flagged as potential outliers
that should be investigated more carefully.

The arguments to \code{makeDataReport()} can be used to modify the
contents and the look of the data report according to the user's
needs. The most commonly used arguments are summarized in
Table~\ref{table.cleanFormals} and they are grouped according to the
part of the data assesment and report generation they influence. In
order to understand this distinction, a glimpse of the inner structure
of \code{makeDataReport()} is shown in
Figure~\ref{figure:cleanStructure}. Below, we present a few examples
on how to use the arguments from Table \ref{table.cleanFormals} to
influence the output of a \code{makeDataReport()} call.

@@

\begin{Schunk}
\begin{Sinput}
R> makeDataReport(toyData, output="html", render=FALSE,
+    openResult=FALSE, replace=TRUE)
\end{Sinput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
R> makeDataReport(toyData, preChecks = c("isKey", "isSupported"), 
+    replace=TRUE)
\end{Sinput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
R> makeDataReport(toyData, visuals = setVisuals(all = "basicVisual"), 
+    replace=TRUE)
\end{Sinput}
\end{Schunk}

In summary, and as indicated in Figure~\ref{figure:cleanStructure}, there are two stages
where \code{makeDataReport()} applies functions to each of the variables:
\begin{enumerate}
\item In the pre-check stage.
\item As part of the summarize/visualize/check (SVC) steps.
\end{enumerate}
Each of these stages are controllable using appropriate function
arguments in \code{makeDataReport()}, and above we have shown examples
of how to tweak them to modify the data cleaning outputs. However, if
the dataset at hand requires new, additional checks, then more control
is needed a package vignette explains the details of how to modify and
expand the possibilities by producing new summary, visual, and check
functions.

@@

\begin{Schunk}
\begin{Sinput}
R> complexData <- data.frame(complexVar = complex(100, real = 1:100, 
+    imaginary = 3), numericVar = 1:100)
R> makeDataReport(complexData, treatXasY=list(complex = "numeric"), 
+    replace=TRUE)
\end{Sinput}
\end{Schunk}

@@

\begin{figure}[tb]
\begin{center}
\includegraphics[width=7.5cm]{article_vol3-toydataevents.pdf}
\end{center}
\caption{Output from running \code{visualize()} on variable \code{events} from the
\code{toyData} dataset.}
\label{fig:example3}
\end{figure}

@@

\begin{figure}[tb]
\begin{center}
\frame{\includegraphics[width=7.5cm,page=3]{dataMaid_presidentData.pdf}}
\frame{\includegraphics[width=7.5cm,page=4]{dataMaid_presidentData.pdf}}
\end{center}
\caption{The second and third pages of the \code{presidentData} data report. We see that there are two \code{Name} variables in the overview on page 2 and see that these variables are indeed treated as \code{character} variables on page 3, as specified in the \code{makeDataReport} call by use of the \code{treatXasY} argument. Larger versions of the pages can be seen in
  Appendix~\ref{sec:appendix2}.}
\label{fig:bigExampleP23}
\end{figure}

\begin{figure}[tb]
\begin{center}
\frame{\includegraphics[width=7.5cm,page=5]{dataMaid_presidentData.pdf}}
\frame{\includegraphics[width=7.5cm,page=6]{dataMaid_presidentData.pdf}}
\end{center}
\caption{Part of the data overview report generated for the \code{presidentData} dataset. Here, we see pages 4 and 5 from the \textit{Variable list}. Larger versions of the pages can be seen in
  Appendix~\ref{sec:appendix2}.}
\label{fig:bigExampleP45}
\end{figure}

We will now put the bits and pieces from above together and show how \code{makeDataReport()} can be used on a less artificial dataset to create a useful overview report and how the interactive tools can subsequently be used to assist the actual data cleaning process. More specifically, we will create a report describing the \code{presidentData} dataset, which is available in \pkg{dataMaid} and use the information from this report to clean up the data. \code{presidentData} is a sligtly mutilated dataset with information about the 45 first US presidents, but with a few common data issues and a blind passenger. The dataset contains one observation per president and has the following variables:
\begin{description}
\item[\code{lastName}] The last name of the president.
\item[\code{firstName}] The first name of the president.
\item[\code{orderOfPresidency}] The number in the order of presidents.
\item[\code{birthday}] The birthday of the president.
\item[\code{stateOfBirth}] The state in which the president was born.
\item[\code{assassinationAttempt}] Was there an assassination attempt on the president?
\item[\code{sex}] The sex of the president.
\item[\code{ethnicity}] The ethnicity of the president.
\item[\code{presidencyYears}] The duration of the presidency.
\item[\code{ageAtInauguration}] The age of the president at inauguration.
\item[\code{favoriteNumber}] The favourite number of the president (fictional).
\end{description}

\begin{Schunk}
\begin{Sinput}
R> data(presidentData)
R> head(presidentData)
\end{Sinput}
\begin{Soutput}
    lastName firstName orderOfPresidency   birthday  stateOfBirth
1 Washington    George                 1 1732-02-22      Virginia
2      Adams      John                 2 1735-10-30 Massachusetts
3  Jefferson    Thomas                 3 1743-04-13      Virginia
4    Madison     James                 4 1751-03-16      Virginia
5     Monroe     James                 5 1758-04-28      Virginia
6      Adams      John                 6 1767-07-11 Massachusetts
  assassinationAttempt  sex ethnicity presidencyYears
1                    0 Male Caucasian               7
2                    0 Male Caucasian               3
3                    0 Male Caucasian               8
4                    0 Male Caucasian               8
5                    0 Male Caucasian               8
6                    0 Male Caucasian               4
  ageAtInauguration favoriteNumber
1                57    3+0.000000i
2                61    4+0.000000i
3                57    0+1.414214i
4                57   10+0.000000i
5                58    3+0.000000i
6                57    9+0.000000i
\end{Soutput}
\end{Schunk}

We discuss the results of a data overview report generated for this dataset below, but first there are a few special features of the dataset and wishes for the data report that require us to customize it using some of the arguments for \code{makeDataReport}. We have the following points of interest:
\begin{enumerate}
\item A couple of variables are used to store names, namely \code{lastName} and \code{firstName}. In order to use special bibliographical analysis tools on these, and only these, variables, it might be convenient to assign them a special class. Therefore, these variables have been set to have class \code{Name} by use of the base \proglang{R} function \code{class()}. When we wish to make a data report for the dataset, we have to tell \code{makeDataReport()} how to handle such \code{Name}-type variables, as they are not among the supported variable types mentioned in the above. This can be done using the \code{treatXasY} argument.
\item We use the \code{character} class for a few categorical variables that have a lot of different levels, but where it is not a data mistake that these levels each generally have very few observations, e.g. the variable \code{stateOfBirth}. Therefore, we would like to disable the \code{identifyLoners} check (which flags variables with <6 observations in any of the levels) for this variable type. This can be done using the \code{checks} argument.
\item We would like the report to be called "Dirty president data" in order to reflect that it is, indeed, a report concerning dirty data about presidents.
\end{enumerate}
Incorporating these three customizations, we can load the data and generate a report by calling:

\begin{Schunk}
\begin{Sinput}
R> makeDataReport(presidentData, replace = TRUE, 
+    treatXasY = list("Name" = "character"),
+    checks = setChecks(character = 
+    defaultCharacterChecks(remove = "identifyLoners")),
+    reportTitle = "Dirty president data")
\end{Sinput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
R> presidentData$lastName[presidentData$lastName == " Truman"] <- "Truman"
R> presidentData$stateOfBirth[presidentData$stateOfBirth == "New york"] <- 
+    "New York"
R> presidentData$assassinationAttempt <- 
+    factor(presidentData$assassinationAttempt)
R> presidentData$ageAtInauguration <- 
+    as.numeric(presidentData$ageAtInauguration)
\end{Sinput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
R> presidentData$ageAtInauguration <- 
+    as.numeric(as.character(presidentData$ageAtInauguration))
\end{Sinput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
R> birthdayOutlierVal <- 
+    identifyOutliers(presidentData$birthday)$problemValues
\end{Sinput}
\end{Schunk}

@@

\subsection[Incorporating dataMaid in an automated workflow]{Incorporating \pkg{dataMaid} in automated workflow}
The default settings of \code{makeDataReport()} have been set to facilitate easy and quick data report generation, but unfortunately, this also means that it is not ideal for a more programming-oriented workflow, where the function might not be called by a human. For instance, one might be interested in automatically running \code{makeDataReport()} on all datasets received from a certain client, perhaps via email or through a web upload, and returning a data report for them to inspect and comment before ever looking at the data. In this scenario, there are a few issues with the standard data report:
\begin{enumerate}
\item After rendering, the report is automatically opened. This is not very useful, if the processes are supposed to run in the background.
\item The report generation writes messages to the console while producing and rendering the report.
\item Unless specifically told otherwise, every report created for the same dataset (or different datasets with the same storage name in \proglang{R}) will have the same file name.
\end{enumerate}
Please note that the data report does contain information about who, when and how concerning its generation, so even though the default choices for file names do not make it easy to tell different reports for the same dataset apart, it should be rather easy when inspecting the report manually. These three problems can easily be solved by use of the arguments of \code{makeDataReport()}. Whether or not the outputted file is opened can be controlled through the argument \code{open}. How much information is printed in the console can be adjusted by using the argument \code{quiet}. And convenintely introducing small alterations of the file names can be obtained by use of the \code{vol} argument. For instance, we can make a data report for \code{toyData} that is not opened automatically, produces no output to the console and includes the date and time of its creation in the file name:

\begin{Schunk}
\begin{Sinput}
R> makeDataReport(toyData, open = FALSE, quiet = "silent",
+    vol = paste("_", format(Sys.time(), "%m-%d%-%y_%H.%M"), sep = ""))
\end{Sinput}
\end{Schunk}

### 2017-11-09 latex/dataMaid_toyData.Rmd
Updated paper

---
dataMaid: yes
title: toyData
subtitle: "Autogenerated data summary from dataMaid"
date: 2017-11-08 12:08:33
output: pdf_document
documentclass: report
header-includes:
  - \renewcommand{\chaptername}{Part}
  - \newcommand{\fullline}{\noindent\makebox[\linewidth]{\rule{\textwidth}{0.4pt}}}
---
```{r , echo=FALSE, include=FALSE, warning=FALSE, message=FALSE, error=FALSE}
library(ggplot2)
library(pander)
```
# Data report overview
The dataset examined has the following dimensions:

@@

 *  Report creation time: Wed Nov 08 2017 12:08:33

>>>>

---
dataMaid: yes
title: toyData
subtitle: "Autogenerated data summary from dataMaid"
date: 2017-11-09 02:44:43
output: pdf_document
documentclass: report
header-includes:
  - \renewcommand{\chaptername}{Part}
  - \newcommand{\fullline}{\noindent\makebox[\linewidth]{\rule{\textwidth}{0.4pt}}}
---
```{r , echo=FALSE, include=FALSE, warning=FALSE, message=FALSE, error=FALSE}
library(ggplot2)
library(pander)
```
# Data report overview
The dataset examined has the following dimensions:

@@

 *  Report creation time: Thu Nov 09 2017 02:44:43
